Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-05-16 doc/GMP2015/revision_2/response_to_reviewers.txt

Response to Reviewers
=====================

Reviewer #1
-----------
Overall I think this paper is timely and interesting.  It is very nice that the authors have included a number of open questions to emphasize that there is more work to be done in this area.  The notation and terminology is rather dense, however, this is probably unavoidable given the intricacy of the questions being addressed.

I don't think the abstract is significantly improved from the original version; there is still a lot of area-specific terminology that makes it hard to understand what the paper is about.  Perhaps a simple motivating example explained in detail over a couple sentences would help, especially explaining why one would want to model a particular ``quasi-uniform, self-similar, layered material structure'' using a decomposition of geometric constraint systems.

To that end, I would also suggest bringing the definition of geometric constraint system out of Appendix A and putting it right at the beginning of section 2, or perhaps section 2.1.

This line at the end of the abstract and at the end of section 2.1 should be modified for the final version of the paper (and applicable files should be included accordingly): ``We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.''

This paper seems essentially ready for publication.

Response:
- Abstract had substantial rework. While we don't give a complete example in the abstract, we do mention which materials we model in this paper. In the introduction, we do explain much more clearly why we would want to use qusecs.
- We moved the definitions into the body of the paper.
- We provide a link where the video and code will be made available. Currently a work in progress, but we are committed to providing a well-documented, open source, highly functional application. The video will be made sooner.

Reviewer #2
-----------
This paper proposes a recursive decomposition planning algorithm to model a material layer as a solution of a 2D geometric constraint system. While the work seems important in material modeling, the topic is a kind of out of focus of GMP audiences. Besides, I'm not quite satisfied with the organization and structure of the paper, and I feel it is not suitable for publication in current form.

1. The paper is far out of a 10 pages limit.
2. The basis concepts should be better put in the main context, but was put in an appendix (Appendix A).
3. The main algorithm in Section 4 should be stood out seperately and explained in details.
4. The examples should be enumerated.
5. The texts contain various formats and styles: Captioned, Bold fonts, colored, underlined, .... This unusual in scientific papers.

Response:
1. We are now submitting to the main journal.
2. We moved preliminaries into the main body of the paper (section 2).
3. We moved the algorithm into its own subsection and revised the proof to be more clear.
4. ...
5. We have addressed this. We have reformatted our unusual headings and such.

Further Changes
===============

Here we discuss additional changes the reviewers should be aware of.

One of the main issues with our paper was that it would not be a good fit for GMP. After much discussion, we decided that we agree and are now submitting the paper for consideration to be published in the standard journal (see next section in this file). The changes made in this revision are discussed below.

Title: Because of the substantial changes to the paper, we decided the title no longer made sense in the context of the revised paper. As such, we have changed it to `Optimal Decomposition and Recombination of Isostatic Geometric Constraint Systems for Designing Layered Materials'. We believe this also makes the intent of our paper much more clear.

Throughout: Many minor corrections to spelling, grammar, and punctuation. Additionally, many sentences were rephrased/restructured to improve clarity. Also, because our paper is no longer being considered for the special issue and does not have to adhere to the page limit, we have reformatted parts of the paper to use more space to improve readability and conform to standard formatting (addressing concerns of reviewer 2).

Two concepts were further fleshed out and discussed in this iteration: (a) The comparison of our canonical DRP to the previously studied cluster-minimal DRP, and (b) the extension of our theory to a larger class of (hyper)graphs, formally those with an underlying abstract rigidity matroid in which independence corresponds to some type of sparsity, and maximal independence (rigidity) is a tightness condition.. This is reflected throughout the paper, but is especially prevalent in (1), (2), (3), and (8).

Abstract: Substantially altered for clarity. Similar content.

(1) Introduction: Entirely rewritten. A major complaint was that it was difficult to get a big picture understanding of the paper without reading the entire thing. We believe we addressed this with our revised intro (and abstract). This also includes a thorough explanation of `qusecs' and why we believe this notion is important to define and give a name, directly addressing reviewer criticism. Also, the discussion of previous work was moved into the corresponding sections (2) and (4).

(2) Preliminaries: Moved back in the definitions from the appendix (as per reviewer request). Moved in discussion of previous work on DR-plans from (1).

(3) Canonical DR-plan: Improved the section introduction. Clarified/improved the definition of a canonical DR-plan. There are new figures to demonstrate the concepts; instead of the previous trivial (in the sense that it was tree-decomposable) example of three triangles, we use three $C2\times C3$s intersecting on a single triangle. Also, there is a new figure that corrects the previous, erroneous figure discussing cluster minimality in the previous revision. Revised proofs for clarity, but there were no fundamental changes. Put algorithm in a separate section (improved proof as mentioned) as per reviewer 2's request. Moved out cluster minimality conjecture to the conclusion. Added a discussion on the overconstrained problem and NP-hardness (3.3).

(4) Recombination: Improved the section introduction. Moved in discussion of previous work on recombination from (1). Made the section that followed in the previous revision into a subsection (4.5).

(5) and (6) Application sections: No changes.

(8) Conclusion: Moved in conjectures about cluster minimality from (3). Added new open problem (OP2).

(App. A) Proofs: Revised DRP proofs for clarity, but there were no fundamental changes.

Journal Change
==============

Originally the paper was submitted for consideration of publication in the special issue of CAGD for GMP2015. However, pending our major revision, this was changed to the regular issue of CAGD. We received confirmation from the GMP co-chairs that this paper would now be handled by the CAGD editors-in-chief. Below, we have copied the email we received from the conference chairs, stating the change.

Upon submission, the exact content of this section will also be emailed to editors-in-chief, along with our submission number: XX.

If there are additional questions in regard to this issue, please send an email.

Email
-----
GMP2015 <gmp2015@easychair.org> Wed, Mar 25, 2015 at 7:17 AM
To: Troy Baker <tbaker@cise.ufl.edu>
Dear Troy Baker,

even though your GMP submission will not appear in the conference special
issue, but - after a major revision - in a forthcoming regular issue of
CAGD, you still have the chance to present your work at the GMP 2015
conference, and we would be very glad if you accept this invitation.

More information about the conference can be found here:
http://gmp2015.inf.usi.ch/

Note that you may require a visa to come to Switzerland and that the
application process can take some time. So it is best to start early. More
information about visa can be found here: http://gmp2015.inf.usi.ch/visa

Please inform us about your decision whether or not to come to GMP 2015
and to present your work by sending an e-mail to "gmp2015@usi.ch" before
*April 2, 2015* to help us plan the conference program.

With kind regards

Elaine Cohen & Kai Hormann
[GMP 2015 conference chairs]

### 2015-05-16 doc/GMP2015/revision_2/appendix_proofs.tex
Revision 2. 24 hours to submission.

For (2), observe that underconstrained subgraphs of isostatic graphs must have density less than $k$. For (3), observe that, given a isostatic graph, a subgraph with density $k$ must also be isostatic. Then, use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Then it is straightforward application of the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.

@@

\begin{proof}
Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
%
In the reverse direction, we know $C$ is either some non-leaf node (isostatic by definition of a DR-plan) or $G$ itself (isostatic by definition of the problem). Thus, $C_i\cup C_j=C$ is isostatic.
\end{proof}

@@

\begin{observation}\label{observation:no_edges_between_diff}
Let us say $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
\end{observation}

\begin{proof}
Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}, shows that $C_i\cup C_j$ must equal the entire parent $C$.
\end{proof}

@@

The alternative phrasing is due to Lemma \ref{lemma:combined_lemma}, point \ref{lemma:wc_intersection_is_C}.

Let us say $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$). Furthermore, we have the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ which are not empty sets.

Then we assume that there is some third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

For convenience, it is implied that all graphs are the induced graph on $C$.

@@

    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they were not vertex-maximal. These are the graphs $\inducedOnC{R'_i\cup R_j\cup D_{i,j}}$ and $\inducedOnC{R_i\cup R'_j\cup D_{i,j}}$ respectively.

@@

\subsubsection{Proof of Remark \ref{rem:bodypin_is_barjoint}}
\begin{proof}
    We can replace each body that has only one pin by a single vertex. A body with 2 pins can be replaced by an edge. In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices. When finding a DR-plan, we treat each body as trivial, so they become the leaves of our plan. The optimal recombination problem and approach of Section \ref{sec:recomb} are unchanged. The optimal completion via the optimal modification problem in Section \ref{sec:recomb} now has an additional constraint that all edges in the 2-tree representation of the bodies must be removed together, not individually.
\end{proof}

@@

\subsubsection{Proof of Theorem \ref{thm:1dofcase}}
\begin{proof}
    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ will then be a vertex maximal proper 1-dof sub-graph of $G$.

    To make the graph isostatic, we need only add one body and pin it to 2 other bodies. Doing so will cause $G$ to become $(3,3)$-tight. We choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children. From Observation \ref{rem:1dofcanon}, we know that the children can only be joined by a single pin or a sub-graph. We pin the new body to bodies in two separate children. Doing so will ensure that all children of $b$ will have 1-dof and all ancestors of $b$ (including $b$) will now be isostatic.

@@

    Such a pinning covers all possible ways of adding a new body. Assume we add a new body $b$ to our graph and pin it to $b_i$ and $b_j$ to make it isostatic. Then, there will be some lowest 1-dof node $v$ in $T$ such that $b_i$ and $b_j$ appear in $v$. Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning $b$ to $b_i$ and $b_j$.

    For each node $b$, we assign a size of the $T_b$ denoted $|T_b|$. $|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$. We are looking for $b$ that minimizes $|T_b|$. Denote the sub-tree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$. Note that $fanin(b)= nl(T^b)$ because no descendant of $b$ is isostatic. Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$. All other nodes are not isostatic and so have no fanin.

    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$. In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic. Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$, excluding $b'$. Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan. The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$. Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.

    The node we need to pin will always be the deepest nontrivial node of some path in $T$. Suppose we chose to pin a node $b$ that has a nontrivial child $v$. Then, $fanin_b(b) = nl(T^b) = nl(T^v) + n$, where $n$ is essentially the number of leaves between $b$ and $v$. If we had instead chosen to pin $v$, then $fanin_v(b) = nl(T^b) - nl(T^{b'}) + 1 \leq fanin_b(v)$. And for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $|T_v| \leq |T_b|$. Thus we only have to check the deepest non-trivial nodes.

    Running this algorithm in a brute force fashion is quadratic in the number of bodies of our given body-pin system.

    For the multi-triangle pin graphs, we can do the same thing except we need to add a single triangle to one of the nodes to cause it to become isostatic.
    % Don't leave a blank line between last paragraph and end!
    % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node. These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1-dof case.
\end{proof}

\subsubsection{Proof of Observation \ref{obs:2dof_case}}
\begin{proof}
    The only difference from the 1-dof case is that now we need to remove 2-dof from our graph. We build a 2-dof DR-plan $T$. Like above, we need to add a body and 2 pins to 2 nodes now to get to isostatic.

    Suppose we pin 2 distinct nodes $v_i$ and $v_j$. Then, there will be some common ancestor $a$ of $v_i$ and $v_j$. Then, in $T_{v_i,v_j}$, $fanin_{v_i,v_j}(a) = nl(T^a)$. However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin_v(a) = nl(T^a) - nl(T^{a'}) +1$ . Thus $fanin_v(a)' \leq fanin_{v_i,v_j}(a)$. All ancestors of $a$ are unchanged. So $|T_v| \leq |T_{v_i,v_j}|$.

    So, we will always end up pinning a single node twice. Hence, we can run the same algorithm as the 1-dof case and just pin twice instead of once.
\end{proof}

\subsubsection{Proof of Observation \ref{obs:algebraic_completion}}
\begin{proof}
    An isostatic graph has 3 parameters that define it (in terms of coming up with a realization). These are the Euclidean motions. A 1-dof graph can be seen as essentially having 4 parameters: the 3 Euclidean motions and whatever degree of freedom it has. A 2-dof has 5 parameters, and so on. These parameters denote the algebraic complexity of giving a realization for the graphs.

    Thus, say we have a $T$ as described in the proof for Remark \ref{thm:1dofcase}. When we pin a node $b$, we will have the same structure as before. Suppose we are looking at an isostatic node $v$ after pinning $b$. Then, the children of $v$ (except one if $v \neq b$) will be 1-dof. The complexity for that node is then just that of solving each of its children. In general then, the number of parameters for that node will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.

    Now if we want to minimize the algebraic complexity, we need to minimize the maximum $np(v)$ for any node $v$. In this case, we cannot always just choose the node furthest down the tree to pin because it could have many children. So we will have to try pinning all nodes to see which gives the lowest algebraic complexity. This is still quadratic for the 1-dof case.

    For the 2-dof case, there are more cases to consider. If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$. If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof. Note that solving $v'$ will result in solving $v$. Then, we need to consider nodes above and including $v'$ in our complexity. $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$. The only remaining case will be if we pin two nodes that are not descendant/ancestor. the only change from the previous case will be that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$. For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.

    Like the 1-dof case, we again cannot just choose the nodes deepest in the tree to pin. However, we cannot also assume pinning one node twice will give us the best algebraic complexity. Hence, we will need to check each pair of node to pin. This makes our brute-force algorithm $O(b^3)$, where $b$ is the number of bodies.
\end{proof}

\subsection{Proofs from Section \ref{sec:pinnedline}}
\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc} --- For 2-dimensional pinned line incidence graphs}
\label{sec:appendix_pinned}

\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
We use the same notations as in the original proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, in the last section.
For convenience, it is implied that all graphs are the induced graph on $C$.

@@

Then we analyze all the possible cases for $C_k$.
\begin{itemize}
    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_k\cup C_i = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
    By Lemma~\ref{lem:pinned_union_intersection}, it must be underconstrained,
    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.

@@

All remaining cases are similar to the original proof.
\end{proof}

>>>>

For the next parts, we use the quantity $d(G)=2|V|-|E|$, which we call \dfn{density}.
For (2), observe that underconstrained subgraphs of isostatic graphs must have density less than 3.
For (3), observe that, given an isostatic graph, a subgraph with density $3$ must also be isostatic. Then, use the fact that, by definition, density of $F_i$ denoted $d(F_i)=3$ and $d(F_j)=3$. Then it is straightforward application of the inclusion-exclusion $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.

@@

\begin{proof}
Assume $C_i\cup C_j \neq C$. This would contradict the proper vertex-maximality of  $C_i,C_j$.
%
In the reverse direction, we know $C$ is either a non-leaf node (isostatic by definition of a DR-plan) or $G$ itself (isostatic by definition of the problem). Thus, $C_i\cup C_j=C$ is isostatic.
\end{proof}

@@

\begin{observation}\label{observation:no_edges_between_diff}
Take $R_i=C\setminus C_i$ and $R_j=C\setminus C_j$. If $C_i\cup C_j$ is isostatic, then there can be no edges in $C$ between the vertices of $R_i$ and $R_j$.
\end{observation}

\begin{proof}
Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, shows that $C_i\cup C_j$ must equal the parent graph $C$.
\end{proof}

@@

The alternative phrasing is as in Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}.

Take $R_i=C\setminus C_i$, $R_j=C\setminus C_j$, and $D_{i,j}=C_i\cap C_j=(C\setminus R_i)\setminus R_j$ (note that $R_j\subset C_i$, $R_i\subset C_j$ and $D_{i,j}\subset C_i,C_j$). Furthermore, take the proper subgraphs $R'_i\subset R_i$, $R'_j\subset R_j$, and $D'_{i,j}\subset D_{i,j}$ that are non-empty.

Assume that there is a third isostatic vertex-maximal proper subgraph $C_k$ (with $C'_k=C\setminus C_k$). There are $3\times 3\times 3 = 27$ possible cases for what this subgraph could be.

Without loss of generality, all graphs are the induced subgraphs of $C$.

@@

    \item 2 cases: $C_k$ cannot contain $C_i$ or $C_j$ as proper subgraphs, or else they are not vertex-maximal. These are the graphs $\inducedOnC{R'_i\cup R_j\cup D_{i,j}}$ and $\inducedOnC{R_i\cup R'_j\cup D_{i,j}}$ respectively.

@@

\subsubsection{Proof of Remark \ref{rem:bodypin_is_barjoint}}
\begin{proof}
    We can replace each body that has only one pin by a single vertex. A body with 2 pins can be replaced by an edge. In general, a body with $n$ pins can be replaced by a 2-tree on $n$ vertices. When finding a DR-plan, we treat each body as trivial, so they become the leaves of the DR-plan. The optimal recombination problem and approach of Section \ref{sec:recomb} are unchanged. The optimal completion via the optimal modification problem in Section \ref{sec:recomb} now has an additional restriction that all edges in the 2-tree representation of the bodies must be removed together, not individually.
\end{proof}

@@

\subsubsection{Proof of Theorem \ref{thm:1dofcase}}
\begin{proof}
    Suppose we are given a body-pin graph and its corresponding body-bar graph $G$ and have obtained the 1-dof DR-plan $T$. Each node of $T$ is then a vertex-maximal proper 1-dof subgraph of $G$.

    To make the graph isostatic, we need only add one body and pin it to 2 other bodies. Doing so will cause $G$ to become $(3,3)$-tight.

We adopt the following algorithm. Choose the 2 bodies to pin to by choosing a node $b$ in $T$ and looking at its children. From Observation \ref{rem:1dofcanon}, we know that the children can only share  a single pin or a sub-graph. Pin the new body to bodies in two separate children. Doing so will ensure that all children of $b$ will have 1-dof and all ancestors of $b$ (including $b$) will now be isostatic.

@@

    Such a pinning covers all possible ways of adding a new body. Assume a new body $b$ is added to the input  graph and pin it to $b_i$ and $b_j$ to make it isostatic. Then, there is a lowest 1-dof node $v$ in $T$ such that $b_i$ and $b_j$ appear in $v$. Thus, pinning $v$ in the manner described yields an equivalent isostatic DR-plan to pinning $b$ to $b_i$ and $b_j$.

    For each node $b$, assign a size of the $T_b$ denoted $|T_b|$. $|T_b| = \displaystyle\max_{v \in T_b} fanin(v)$. We are looking for $b$ that minimizes $|T_b|$. Denote the sub-tree of $T$ rooted at $v$ by $T^v$ and the number of leaves in a tree $T$ by $nl(T)$. Note that $fanin(b)= nl(T^b)$ because no descendant of $b$ is isostatic. Similarly, for any ancestor $w$ of $b$, $fanin(w) = nl(T^w)-nl(T^{b'})+1$, where $b'$ is the child leading to $b$. All other nodes are not isostatic and hence do not appear in the isostatic DR-plan.
    % Then, we can form a valid isostatic DR-plan $T_b$ from $T$. In $T_b$, $b$'s children are now all of the leaf nodes of the subtree rooted at $b$ because no child of $b$ in $T$ is isostatic. Similarly, for any other node $w$ that is an ancestor of $b$, $w$'s children will be the node that leads to $b$, denoted $b'$, along with all of the other leaf nodes in the tree rooted at $w$, excluding $b'$. Then, for any node $b$ that we choose, $T_b$ is a valid DR-plan. The size of $T_b$ will just be the maximum fanin of all nodes in $T_b$. Thus, if we want to minimize the size of our DR-plan, we simply need to take the $b$ that has the $T_b$ of smallest size.

    The node to be pinned is always the the deepest nontrivial node of some path in $T$. Suppose a node $b$ is pinned that has a nontrivial child $v$. Then, $fanin_b(b) = nl(T^b) = nl(T^v) + n$, where $n$ is essentially the number of leaves between $b$ and $v$. If we had instead chosen to pin $v$, then $fanin_v(b) = nl(T^b) - nl(T^{b'}) + 1 \leq fanin_b(v)$. And for each ancestor $w$ of $b$, $fanin(w)$ is unchanged, meaning $|T_v| \leq |T_b|$. Thus we only have to check the deepest non-trivial nodes.

    Running the above algorithm brute force  gives running time quadratic in the number of bodies of the given body-pin system.

    For the multi-triangle pin graphs, we can do the same thing, except we need to add a single triangle to one of the nodes to cause it to become isostatic.
    % Don't leave a blank line between last paragraph and end!
    % For the 2-dof case, we can do something very similar, except instead of pinning a single body 2 times to a node, we can pin another body 2 times to a node. These can be the same node, and if it is the same node, we can find a wellconstrained DR-plan in quadratic time, we would just be doing the same thing as the 1-dof case.
\end{proof}

\subsubsection{Proof of Observation \ref{obs:2dof_case}}
\begin{proof}
    The only difference from the 1-dof case is that now we need to remove
2-dofs from our graph. Start with a 2-dof DR-plan $T$. Like in the previous proof, we need to add a body and 2 pins to 2 nodes to obtain an isostatic DR-plan.

    Suppose we pin 2 distinct nodes $v_i$ and $v_j$. Then, there must exist a common ancestor $a$ of $v_i$ and $v_j$. Then, in $T_{v_i,v_j}$, $fanin_{v_i,v_j}(a) = nl(T^a)$. However, if we chose to pin one of $v_i$ and $v_j$ twice, then $fanin_v(a) = nl(T^a) - nl(T^{a'}) +1$ . Thus $fanin_v(a)' \leq fanin_{v_i,v_j}(a)$. All ancestors of $a$ are unchanged. So $|T_v| \leq |T_{v_i,v_j}|$.

    Thus the only choice is to pin a single node twice. Hence, we can run the same algorithm as the 1-dof case and simply pin twice instead of once.
\end{proof}

\subsubsection{Proof of Observation \ref{obs:algebraic_completion}}
\begin{proof}
    An isostatic graph has 3 parameters that define its position and orientation. These are the Euclidean motions. A 1-dof graph has  4 parameters: the 3 Euclidean motions and a dof parameter. A 2-dof has 5 parameters. The number of parameters roughly correlates with the algebraic complexity of obtaining a realization.

    Thus, starting with a $T$ as described in the proof for Remark \ref{thm:1dofcase}: when a node $b$ is pinned, the same structure is preserved as before. Suppose  $v$ is an isostatic node after pinning $b$. Then, the children of $v$ (except one if $v \neq b$) have 1-dof. The realization complexity for $v$ is simply that of realizing each of its children. In general, the number of parameters for $v$  will be $np(v) = 4nc_1(v)+3$, if $v \neq b$ and $np(b) = 4nc_1(b)$, where $nc_k(v)$ is the number of $k$-dof children of $v$.

    Minimizing the algebraic complexity  requires minimizing the maximum $np(v)$ for any node $v$. In this case, it is not possible to always choose to pin a node closest to a leaf in the tree, because it could have high fan-in. So we  try
brute force by pinning all nodes to  pick the one with the lowest algebraic complexity. This algorithm is still quadratic for the 1-dof case.

    For the 2-dof situation, there are more cases to consider. If we pin the same node twice as above, we have $np(v) = 5nc_2(v)+3$ for any ancestor $v \neq b$ and $np(b) = 5nc_2(b)$. If we pin a node $v$ and one of its ancestors $v'$, then any nodes between $v'$ and $v$ will be 1-dof, any nodes above $v'$ will be isostatic, and nodes below $v$ will be 2-dof. Note that solving or realizing $v'$ will also realize $v$.  Next, we need to consider nodes above and including $v'$ in our complexity: $np(v') = 5nc_2(v') + 4$ and $np(a) = 3 + 5nc_2(a)$ for $a$ an ancestor of $v'$.

The only remaining case is pinning two nodes that are incomparable, i.e., do not have a descendant/ancestor relationship. The only change from the previous case is that for the lowest common ancestor of the nodes $v'$, $np(v') = 2*4+5nc_2(v')$. For any ancestor $a$ of $v'$, we still have $np(a) = 3 + 5nc_2(a)$.

    Like the 1-dof case, we again cannot simply choose the nodes deepest in the tree to pin. However, neither can we assume pinning one node twice will give us the best algebraic complexity. Hence, we will need to check each pair of nodes to pin. This makes our brute-force algorithm $O(b^3)$, where $b$ is the number of bodies.
\end{proof}

\subsection{Proofs from Section \ref{sec:pinnedline}}
\subsubsection{Proof of Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc} --- for 2-dimensional pinned line incidence graphs}
\label{sec:appendix_pinned}

\begin{proof} %[Proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for 2D pinned line incidence graphs]
We use the same notation as in the original proof of  Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, given above.
Without loss of generality, all graphs are the induced graphs on $C$.

@@

We analyze all the possible cases for $C_k$.
\begin{itemize}
    \item 1 case: $C_k=\inducedOnC{R'_i\cup R'_j\cup D_{i,j}}$ is not possible. Since $C_k\cup C_i = \inducedOnC{R'_i\cup R_j\cup D_{i,j}}\neq C$ we have from Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_is_C}, that $C_k\cup C_i$ cannot be isostatic.
    By Lemma~\ref{lem:pinned_union_intersection}, it must be underconstrained,
    so one of $C_i$ and $C_k$ must be an edge, contradicting the assumption that both $C_i$ and $C_k$ are isostatic.

@@

All remaining cases are similar to the original proof for 2D bar-joint graphs.
\end{proof}
